Synthesizing Quotient Lenses

### 2017-10-13 papers/QuotientLensSynth/main.tex
Comments

However, the constraints that are imposed by the lens laws are sometimes too
strict. For instance, the programmer may wish to ignore updates to the data
that only change the order of fields, or add whitespace characters that in
reality are not part of the data. Foster et al therefore proposed
\textit{quotient lenses} ~\cite{quotient} as a solution to this problem
Quotient lenses, which are implemented as a refinement of the bidirectional
string processing language Boomerang, enable a programmer to write lenses which
transform data that is quotiented out by an equivalence relation, so that
records which are equivalent to each other modulo this relation are considered
the same by the underlying lens.

@@

Unfortunately, quotient lenses are sometimes difficult to program in Boomerang,
firstly because the programmer has to give a specification for the data as well
as give a separate specification for the equivalence relations defined on the
data. For example, suppose that the programmer wants to program the \bibtex{} to
EndNote transformation in such a way that the program allows for any of two
representations of the authors name (e.g. ``Conway, J. H." and ``Conway J. H"),
ignores the order in which the author and title fields occur in each record,
and allows for arbitrary amounts of whitespace in between the author and title
fields. These rules allow for these two \bibtex{} citation records to be
considered the same,

@@

to derive a quotient lens that does the \bibtex{} to EndNote transformation.

@@

\subsection{Normal Forms of QRE Lenses}
Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
\Leftrightarrow c'$ be such that 
\begin{align*}
\llbracket q \rrbracket.get &= \ell \circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \ell^{-1} \circ
\canonizer(c')
\end{align*}
for some bijective lens $\ell$. We now provide a proof that all QRE lenses are
of this form.
\begin{theorem}\label{normal form}
If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
\end{theorem}
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
  $\ell :
  R \Leftrightarrow S$. Then
  \begin{align*}
  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
  \end{align*}
  \item
  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ \canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
  (\canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}

@@

\bcp{Make sure to list/discuss:
  \begin{itemize}
  \item Seminar papers on lenses (to which we can refer readers for further
  reading about connections to view updates, etc.) and quotient lenses.
  \item Discuss what other lens-like programming languages (besides
  boomerang) do about quotienting.  E.g., Augeas, XSugar, ...
  \item How quotient lenses are used in Boomerang.
  \item Search for papers that cite the quotient lens paper and see if they
  extend the ideas there.
  \end{itemize}
}

>>>>

However, the constraints that are imposed by the lens laws are sometimes too
strict. For instance, the programmer may wish to ignore updates to the data
that only change the order of fields, or add whitespace characters that in
reality are not part of the data. Foster {\em et al.\relax} therefore proposed
\textit{quotient lenses} ~\cite{quotient} as a solution to this problem
Quotient lenses, which are implemented as a refinement of the bidirectional
string processing language Boomerang, enable a programmer to write lenses which
transform data that is quotiented out by an equivalence relation, so that
records which are equivalent to each other modulo this relation are considered
the same by the underlying lens.

@@

Unfortunately, quotient lenses are sometimes difficult to program in Boomerang,
firstly because the programmer has to give a specification for the data as well
as give a separate specification for the equivalence relations defined on the
data. For example, suppose that the programmer wants to program the \bibtex{} to
EndNote transformation in such a way that the program allows for any of two
representations of the authors name (e.g. ``Conway, J. H.'' and ``Conway
J. H''\bcp{``J. H. Conway''??}), 
ignores the order in which the author and title fields occur in each record,
and allows for arbitrary amounts of whitespace in between the author and title
fields. These rules allow for these two \bibtex{} citation records to be
considered the same,\bcp{I thought
  we were going to allow newlines within titles and authors...?}

@@

to derive a quotient lens that does the \bibtex{} to EndNote transformation.
\bcp{What about examples??}

@@

\subsection{Normal Forms of QRE Lenses}
Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
\Leftrightarrow c'$ be such that 
\begin{align*}
\llbracket q \rrbracket.get &= \ell \circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \ell^{-1} \circ
\canonizer(c')
\end{align*}
for some bijective lens $\ell$. We now provide a proof that all QRE lenses are
of this form.\bcp{Maybe some cases of the proof can be elided in the short
  version?} 
\begin{theorem}\label{normal form}
If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
\canonizer(c')
\end{align*}
\end{theorem}
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
  $\ell :
  R \Leftrightarrow S$. Then
  \begin{align*}
  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
  \end{align*}
  \item
  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ \canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
  (\canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}

@@

\bcp{Make sure to list/discuss:
  \begin{itemize}
  \item Seminar papers on lenses (to which we can refer readers for further
  reading about connections to view updates, etc.) and quotient lenses.
  \item Discuss what other lens-like programming languages (besides
  boomerang) do about quotienting.  E.g., Augeas, XSugar, ...
  \item How quotient lenses are used in Boomerang.
  \item Search for papers that cite the quotient lens paper and see if they
  extend the ideas there.
  \item Search for theoretical papers involving equivalence relations over
  regular languages. 
  \end{itemize}
}
